---
title: "HW6: Text Mining"
author: "Chenxi Zhu, Teresa Chang, and Tiffany Luo"
date: "`r Sys.Date()`"
output: prettydoc::html_pretty
editor_options: 
  markdown: 
    wrap: 80
  chunk_output_type: inline
---

# 1. Introduction

In this project, we aim to explore the textual characteristics of the 20 most
popular books obtained from the Gutenberg Project, focusing on Mary Shelley's
timeless classic Frankenstein. By employing text mining techniques, we aim to
delve into the intricate pattern of the narrative by using text analysis tools
in R, such as data preprocessing, term document matrix creation, document term
matrix analysis, text clustering, and sentiment analysis. Besides looking at
word frequencies, we investigate the sentiment distribution across various
chapters and novels, unveiling the emotional components and tones in the
narratives. This analysis hopes to reveal the mystery of linguistic nuances and
thematic terms in Frankenstein with the hope of understanding the Gothic
masterpiece and its cultural and historical contexts better. It can be extended
to other literary genres, providing a comprehensive understanding of
genre-specific linguistic nuances and diverse cultural landscapes.

# 2. Methods

In this project, text-mining techniques are employed to analyze the book
Frankenstein. The process begins with data preprocessing. The text from the
provided URLs is converted into a corpus, and we clean the datasets by removing
special characters, numbers, punctuation, and common English stop words so that
the remaining can show the frequency of the word uses without being influenced
by the words that don\'t provide helpful information. Additional self-defined
stop words are also eliminated, and stemming or lemmatization is applied for
further refinement. 

-   Stemming is a normalization technique that removes suffixes to reduce words
    to their root. For example, stemming would transform the words \'running\'
    and \'runner\' to the word root of \'run.\'

-   Lemmatization is also a normalization technique. Compared to stemming,
    lemmatization considers the context of the word in the sentence and produces
    a meaningful base form, such as transforming \'better\' to \'good.\'

After that, the term document matrix (TDM) is created, and non-Scrabble
Dictionary words are filtered out, which helps focus analyses on meaningful and
widely recognized terms. TDM is a mathematical representation of a corpus. The
rows represent unique terms, and the columns represent documents. Each cell in
the matrix contains the frequency of a particular term in a specific document,
making it a structured way to analyze the occurrence of words in a text
collection.

Subsequently, a document term matrix (DTM) is generated, similar to a TDM but
with rows and columns switched. The distribution of terms is visualized using
histograms and word clouds, which help visually represent the frequency of words
in a text. The more frequently occurring words are displayed in larger fonts.

Furthermore, data preparation involves filtering out low-frequency terms, and
k-means analysis is performed to cluster terms based on their relative
frequencies. This analysis iteratively assigns data points to clusters and
adjusts the cluster centroids until the clusters are optimized, allowing the
identification of patterns or groups within a dataset. The optimal number of
clusters is determined using the NbClust approach, which considers various
clustering indices, including the Hubert index and the D index, to assess the
quality and stability of clusters for different values of k, helping to choose
the most suitable number of clusters for a given dataset.

Lastly, sentiment analysis is the process of categorizing the emotional tone
expressed in a piece of text, often as positive, negative, or neutral. This
analysis is performed using sentiment lexicons, lists that associate words with
specific information that assign sentiment scores to words. Therefore, multiple
lexicons, including Syuzhet, Bing, AFINN, and NRC, are compared at the end, and
sentiment scores are visualized.

# 3.Results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here, we have a list of 20 most popular books which we obtain from the Gutenberg
project (<https://www.gutenberg.org/ebooks/search/?sort_order=downloads>). These
books are sorted by popularity/number of downloads.

```{r libraries, message=FALSE, warning=FALSE}
library(wordcloud)
library(text)
library(tm)
library(SnowballC)
library(words)
library(NbClust)
library(stringr)
library(dplyr)
library(syuzhet)
```

## 3.1 Load and preprocess the text of the books from www.gutenberg.org.

```{r load-data}
# Define the URLs of the text documents
urls <- c(
  "https://www.gutenberg.org/cache/epub/84/pg84.txt",
  "https://www.gutenberg.org/cache/epub/2701/pg2701.txt",
  "https://www.gutenberg.org/cache/epub/46/pg46.txt",
  "https://www.gutenberg.org/cache/epub/1342/pg1342.txt",
  "https://www.gutenberg.org/cache/epub/1513/pg1513.txt",
  "https://www.gutenberg.org/cache/epub/145/pg145.txt",
  "https://www.gutenberg.org/cache/epub/2641/pg2641.txt",
  "https://www.gutenberg.org/cache/epub/100/pg100.txt",
  "https://www.gutenberg.org/cache/epub/37106/pg37106.txt",
  "https://www.gutenberg.org/cache/epub/16389/pg16389.txt",
  "https://www.gutenberg.org/cache/epub/67979/pg67979.txt",
  "https://www.gutenberg.org/cache/epub/394/pg394.txt",
  "https://www.gutenberg.org/cache/epub/6761/pg6761.txt",
  "https://www.gutenberg.org/cache/epub/2160/pg2160.txt",
  "https://www.gutenberg.org/cache/epub/4085/pg4085.txt",
  "https://www.gutenberg.org/cache/epub/6593/pg6593.txt",
  "https://www.gutenberg.org/cache/epub/5197/pg5197.txt",
  "https://www.gutenberg.org/cache/epub/1259/pg1259.txt",
  "https://www.gutenberg.org/cache/epub/31516/pg31516.txt",
  "https://www.gutenberg.org/cache/epub/11/pg11.txt"
)
```

## 3.2 Data Preprocessing

The first thing we want to do is to convert the text in all of these URLS into a
Corpus.

```{r warning=FALSE, message=FALSE, cache=FALSE}
# Load and preprocess all text documents
myCorpus <- tm::VCorpus(VectorSource(sapply(urls, readLines)))

# Convert everything to lower case
myCorpus <- tm_map(myCorpus, content_transformer(tolower))
```

We'll look at lines 830-900 of the first entry (document) in the corpus, which
can exclude information about the ebook, and the Table of Contents.

```{r warning=FALSE, message=FALSE, cache=FALSE}
cat(content(myCorpus[[1]])[830:900], sep = "\n")
```

Now that we have the data in a corpus, we'll do some data cleaning, by
converting a bunch of special characters (e.g., \*\*@*\*, **/**, **]**, **\$**)
to a space and by removing apostrophes.

```{r warning=FALSE, message=FALSE, cache=FALSE}
#     Defining the toSpace function
toSpace <- content_transformer(function(x, pattern) gsub(pattern, " ", x))
#     Defining the remApostrophe function
remApostrophe <- content_transformer(function(x,pattern) gsub(pattern, "", x))
#     Removing special characters
myCorpus <- tm_map(myCorpus, toSpace, "@")
myCorpus <- tm_map(myCorpus, toSpace, "/")
myCorpus <- tm_map(myCorpus, toSpace, "]")
myCorpus <- tm_map(myCorpus, toSpace, "$")
myCorpus <- tm_map(myCorpus, toSpace, "—")
myCorpus <- tm_map(myCorpus, toSpace, "‐")
myCorpus <- tm_map(myCorpus, toSpace, "”")
myCorpus <- tm_map(myCorpus, toSpace, "‘")
myCorpus <- tm_map(myCorpus, toSpace, "“")
myCorpus <- tm_map(myCorpus, toSpace, "‘")
myCorpus <- tm_map(myCorpus, remApostrophe, "’")
```

We'll then remove numbers and punctuation.

```{r warning=FALSE, message=FALSE, cache=FALSE}
myCorpus <- tm::tm_map(myCorpus, removeNumbers)
myCorpus <- tm_map(myCorpus, removePunctuation)
cat(content(myCorpus[[1]])[830:900], sep = "\n")
```

We look at a list of English stop words (e.g., *a*, *to*) that we can remove
from the documents. Stop words are frequent terms that often don't provide a lot
of useful information.

```{r warning=FALSE, message=FALSE, cache=FALSE}
stopwords("english")
myCorpus <- tm_map(myCorpus, removeWords, stopwords("english"))
cat(content(myCorpus[[1]])[830:900], sep = "\n")
```

We can also remove additional (i.e., self-defined) stop words, such as *ebook*
that appear in the first few lines of the text.

```{r warning=FALSE, message=FALSE, cache=FALSE}
myCorpus <- tm_map(myCorpus, removeWords,c("ebook", "author", "translators", 
                                           "start wwwgutenbergorg", "gutenberg", "chapter", "book", "contents"))
cat(content(myCorpus[[1]])[830:900], sep = "\n")
```

Lastly, We can play around with stemming. This removes common word suffixes and
endings like *es*, *ed*, *ing*, etc. Alternatively, there is lemmatization,
which groups together different inflected forms of the same word.

```{r warning=FALSE, message=FALSE, cache=FALSE, eval=FALSE}
myCorpus <- tm_map(myCorpus, stemDocument)
cat(content(myCorpus[[1]])[830:900], sep = "\n")
```

## 3.3 Term Document Matrix

The next step is to create a term document matrix (TDM). A TDM is a
representation of how frequently different terms (shown in rows) appear in each
of the documents (shown in columns). The transpose of the TDM is the document
term matrix (DTM), where the rows and columns are switched.

```{r warning=FALSE, message=FALSE, cache=FALSE}
tdm <- TermDocumentMatrix(myCorpus)
tm::inspect(tdm)
```

We convert the TDM to a matrix that we call `m`. Each of the rows in `m`
corresponds to each of the unique terms (words) that appears in the documents,
and each of the columns corresponds to each document.

```{r warning=FALSE, message=FALSE, cache=FALSE}
m<- as.matrix(tdm)
dim(m)
rownames(m) <- tdm$dimnames$Terms
colnames(m) <- c("Frankenstein", "Moby Dick", "A Christmas Carol in Prose", "Pride and Prejudice", "Romeo and Juliet", "Middlemarch", "A Room with a View", 
"The Complete Works of William Shakespeare", "Little Women", "The Enchanted April", "The Blue Castle", "Cranford", "The Adventures of Ferdinand Count Fathom", "The Expedition of Humphry Clinker", "The Adventures of Roderick Random", "History of Tom Jones, a Foundling", "My Life — Volume 1", "Twenty years after", "The Eyes Have It", "Alice's Adventures in Wonderland")
head(m)
```

We filter our data to include only those words that are actually in the Scrabble
Dictionary.

```{r warning=FALSE, message=FALSE, cache=FALSE}
dictionary <- as.character(words::words$word)
row_names <- rownames(m)
in_dictionary <- row_names %in% dictionary
remove <- as.character(row_names[!in_dictionary])

#Since the data are so large, if we try to remove all words at once, we get an error. So we will remove them in chunks of 1000.

num_observations <- as.numeric(length(remove))  # Total number of observations
chunk_size <- 1000                              # Number of observations to display at a time

for (i in seq(1, num_observations, chunk_size)) {
  start <- i
  end <- i + chunk_size - 1
  end <- ifelse(end > num_observations, num_observations, end)
  myCorpus <- tm_map(myCorpus, removeWords, remove[start:end])  
}
```

We look at the terms that were dropped in myCorpus -- these are all the words
that aren't in the Scrabble Dictionary

```{r warning=FALSE, message=FALSE, cache=FALSE}
cat(content(myCorpus[[1]])[980:1000], sep = "\n")
```

## 3.4 Document Term Matrix

Now that `myCorpus` has had all the words that aren't in the dictionary removed,
we convert it to a document term matrix (DTM). This is the format that we
actually want - where the documents are rows, and terms are columns.

```{r warning=FALSE, message=FALSE, cache=FALSE}
dtm_cleaned <- DocumentTermMatrix(myCorpus)
tm::inspect(dtm_cleaned)
```

As earlier, we convert the DTM to a matrix.

```{r warning=FALSE, message=FALSE, cache=FALSE}
m <- as.matrix(dtm_cleaned)
dim(m)
colnames(m) <- dtm_cleaned$dimnames$Terms
rownames(m) <- c("Frankenstein", "Moby Dick", "A Christmas Carol in Prose", "Pride and Prejudice", "Romeo and Juliet", "Middlemarch", "A Room with a View", 
"The Complete Works of William Shakespeare", "Little Women", "The Enchanted April", "The Blue Castle", "Cranford", "The Adventures of Ferdinand Count Fathom", "The Expedition of Humphry Clinker", "The Adventures of Roderick Random", "History of Tom Jones, a Foundling", "My Life — Volume 1", "Twenty years after", "The Eyes Have It", "Alice's Adventures in Wonderland")
```

We look at the term distribution. The histogram and tabulation can show us the
distribution of term frequency across all terms. We can see that there are a lot
of terms that appear only a few times, while some others appear thousands of
times. The word cloud shows us the terms that appear the most, such that higher
frequency is indicated by a larger font.

```{r warning=FALSE, message=FALSE, cache=FALSE}
cs <- as.matrix(colSums(m))             #How many times each term appears across all documents (texts)
rownames(cs) <- dtm_cleaned$dimnames$Terms

hist(cs, breaks=100)                    #histograms/tabulations/word cloud of total term appearance 
tab <- as.matrix(table(cs))
wordcloud(myCorpus, min.freq=1000)
```

In the word cloud, we can see that words such as "thou", "well", "time", "lord"
are some of the most often used words in Frankenstein. Since the novel is a
Gothic work of fiction in Romantic era, it makes sense that its language
reflects the literary conventions of that time. For example, "Thou" is an
archaic second-person singular pronoun, which contributes to the somewhat
archaic atmosphere of the novel. The use of the word "lord" could be tied to the
novel's Gothic and Romantic elements. It might be used to represent a sense of
awe, reverence, or to emphasize the power dynamics in the novel.

## 3.5 Text Clustering

In our preparation for cluster analysis, we remove all variables where the
column sum is less than 10,000 (i.e., if the term appears less than 10,000 times
in all documents). We are only doing it here so that we have a reasonable number
of variables to include in the cluster analysis for the sake of this example.

```{r warning=FALSE, message=FALSE, cache=FALSE}
variables_to_remove <- cs < 10000

# Subset matrix frame, excluding those variables
m_subset <- m[, !variables_to_remove]
```

We do some additional data preparation for the cluster analysis.

```{r warning=FALSE, message=FALSE, cache=FALSE}
#Some books are longer, others are shorter. We divide the frequencies by the total number of words (after processing) in each book.
m_fin <- m_subset/rowSums(m)

#Let's scale (normalize) each of the variables (relative frequency)
m_scale <- scale(m_fin)
```

Before performing the k-means analysis on the scaled (normalized) relative
frequencies of words, we need to identify the optimal number of clusters based
on the result of NbClust approach.

```{r warning=FALSE, message=FALSE, cache=FALSE}

#NbClust approach
set.seed(1234)
nc <- NbClust(m_scale, min.nc=2, max.nc=15, method="kmeans", index="all")
#table(nc$Best.n[1,])
par(mfrow=c(1,1)) 
barplot(table(nc$Best.n[1,]),
        xlab="Numer of Clusters", ylab="Number of Criteria",
        main="Number of Clusters Chosen by 26 Criteria")
```

We will use 3 clusters based on the NbClust approach.

```{r warning=FALSE, message=FALSE, cache=FALSE}
k_means_results <- kmeans(m_scale, 3, 30)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
k_means_results$cluster
k_means_results$size
```

Finally, we look at the number of times each of the terms appears in each
cluster. Ultimately, we see that certain terms have different frequencies in the
clusters.

```{r warning=FALSE, message=FALSE, cache=FALSE}
word_totals_by_cluster <- round(aggregate(m_subset, by=list(cluster=k_means_results$cluster), sum),1)

#plot the results
#Decrease font size and rotate x-axis labels vertically
par(cex.axis = 0.7)  # Adjust the font size
par(las = 2)        # Rotate labels vertically

barplot(as.matrix(word_totals_by_cluster[-1]),
        beside = TRUE,
        col = c("blue", "green"),
        legend.text = TRUE,
        args.legend = list(x = "topright"))

# Add labels to the x-axis and y-axis. Here, the first cluster is in blue and the second one is in green.
title(xlab = "Cluster")
title(ylab = "Sum")

# Add a title to the plot
title(main = "Bar Plot of Sums by Group")
```

## 3.6 Sentiment Analysis

A sentiment lexicon, also known as a sentiment dictionary, is a collection of
words or phrases annotated with sentiment polarity information. It associates
each word or phrase with a sentiment score indicating its positive, negative, or
neutral sentiment. Sentiment lexicons are commonly used in sentiment analysis
tasks to determine the sentiment or emotional tone expressed in text.

We use sentiment analysis to examine the sentiment in an entire body of text for
the 20 most popular books. This is often done by aggregating sentiment scores of
all the words in the text. One issue is that for longer texts, the positive and
negative terms often tend to wash each other out. Therefore, shorter texts of a
few sentences or paragraphs work well.

Here, we will use nrc as our sentiment lexicon.We take the first book in our
data, Frankenstein, which is the first row of the matrix `m`. The
`get_nrc_sentiment` command obtains the sentiment score for each word in
Frankenstein that wasn't removed in the cleaning process above.

```{r warning=FALSE, message=FALSE, cache=FALSE}
nrc <- syuzhet::get_sentiment_dictionary(dictionary="nrc")
head(nrc, n=20L)

```

```{r warning=FALSE, message=FALSE, cache=FALSE}
afinn <- syuzhet::get_sentiment_dictionary(dictionary="afinn")
head(afinn, n=20L)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
bing <- syuzhet::get_sentiment_dictionary(dictionary="bing")
head(bing, n=20L)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
syuzhet <- syuzhet::get_sentiment_dictionary(dictionary="syuzhet")
head(syuzhet, n=20L)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
get_nrc_sentiment("gorgeous")
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
Frankenstein <- as.data.frame(m[1,])
Frankenstein$Term <- as.vector(rownames(Frankenstein))
colnames(Frankenstein)[1] = "Term_Frequency"
rownames(Frankenstein) <- 1:nrow(Frankenstein)

nrc_sentiment <- get_nrc_sentiment(Frankenstein$Term)
```

We combine the original data frame and the sentiment counts.

```{r warning=FALSE, message=FALSE, cache=FALSE}
Frankenstein_Sentiment <- cbind(Frankenstein, nrc_sentiment)

```

We multiply the sentiment by the frequency of the term.

```{r warning=FALSE, message=FALSE, cache=FALSE}
head(Frankenstein)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
# Select the columns to be multiplied (last ten columns)
cols_to_multiply <- names(Frankenstein_Sentiment)[3:12]


# Multiply the last ten columns (sentiments) by the first column (Term_Frequency)
Frankenstein_Sentiment[, cols_to_multiply] <- Frankenstein_Sentiment[, cols_to_multiply] * Frankenstein_Sentiment$Term_Frequency
```

We'll see the total prevalence of each sentiment in the text by summing each
column and creating a bar plot

```{r warning=FALSE, message=FALSE, cache=FALSE}
Frankenstein_Sentiment_Total <- t(as.matrix(colSums(Frankenstein_Sentiment[,-1:-2])))
barplot(Frankenstein_Sentiment_Total, las=2, ylab='Count', main='Sentiment Scores')
```

From the plot above, we can see that the number of words with positive
sentiments are more than the number of words with negative sentiments. Note that
We're using Frankenstein as an example for the sentiment analysis, and this
process can be done for each of the books to find out their word sentiment
distribution.

We can also try other lexicons in addition to nrc. Specifically, we will use the
`get_sentiment` command to get the scores for each of the terms using each
dictionary.

```{r warning=FALSE, message=FALSE, cache=FALSE}
Frankenstein$Syuzhet <- as.matrix(get_sentiment(Frankenstein$Term, method="syuzhet"))
hist(Frankenstein$Syuzhet)
Frankenstein$Bing <- as.matrix(get_sentiment(Frankenstein$Term, method="bing"))
hist(Frankenstein$Bing)
Frankenstein$AFINN <- as.matrix(get_sentiment(Frankenstein$Term, method="afinn"))
hist(Frankenstein$AFINN)
Frankenstein$NRC <- as.matrix(get_sentiment(Frankenstein$Term, method="nrc"))   #There are Negative and Positive sentiments in the NRC output above.
hist(Frankenstein$NRC)

```

We can compare the results from the different lexicons by assigning a value of
-1 to all terms that have a negative sentiment, 0 to all terms that have a
neutral sentiment, and 1 to all terms that have a positive sentiment.

```{r warning=FALSE, message=FALSE, cache=FALSE}
sentiment_columns <- Frankenstein[ , 3:6]
sentiment_columns <- data.frame(lapply(sentiment_columns, sign))
sentiment_columns <- data.frame(lapply(sentiment_columns, as.factor))
```

We have the prevalence of *unique* negative, neutral terms (here, we are talking
about *unique* terms because we are not weighing them by how often they appear
in the document).

```{r warning=FALSE, message=FALSE, cache=FALSE}
#Raw frequencies
sapply(sentiment_columns, function(x) if("factor" %in% class(x)) {table(x)})
#Proportions
sapply(sentiment_columns, function(x) if("factor" %in% class(x)) {prop.table(table(x))})
```

We can see that in general, most of the terms are neutral, with about 3-8% of
terms being positive and 4-13% of the terms being negative, depending on which
lexicon we use.

# 4. Discussion

In the word Cloud, the presence of words like "thou," "well," "time," and "lord"
suggests that the language used in that piece of "Frankenstein" is likely
reflective of the time period and literary style in which it was written.
Comparing word clouds from different novels or different chapters can highlight
changes in language use, and potentially emphasizing critical events in the
narrative.

Going beyond word frequency, investigating the sentiment distribution across
different chapters or different novels may help to identify the tone of the
narratives. There might be specific chapters or novels associated with more
positive or negative sentiments. This could provide a understanding of the
novel's themes and plot development.

For future analysis, we can use Frankenstein as an example work of Gothic
literature, and explore similar Gothic literature to find whether there are
specific words associated with the Gothic genre in general. This can help to
identify the key words in the genre, and link with the cultural and historical
context. In addition, we can apply this method to more genres and find their
genre key words.

# Reference

-   Ali, A., & Saad, S. (2016). Term extraction and hierarchy induction method
    based on islamic dictionary. <https://doi.org/10.1109/infrkm.2016.7806345>

-   Text Mining in Data and Information: Unveiling Insights through Data Mining
    -- Gosic. <https://gosic.org/text-mining/>
